---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# frogger

<!-- badges: start -->
[![Project Status: Concept – Minimal or no implementation has been done yet, or the repository is only intended to be a limited example, demo, or proof-of-concept.](https://www.repostatus.org/badges/latest/concept.svg)](https://www.repostatus.org/#concept)
[![R build status](https://github.com/mrc-ide/frogger/workflows/R-CMD-check/badge.svg)](https://github.com/mrc-ide/frogger/actions)
[![codecov.io](https://codecov.io/github/mrc-ide/frogger/coverage.svg?branch=main)](https://codecov.io/github/mrc-ide/frogger?branch=main)
<!-- badges: end -->

Leapfrog is a multistate population projection model for demographic and HIV epidemic estimation.

The name _leapfrog_ is in honor of [Professor](https://iussp.org/en/basia-zaba-1949-2018) Basia [Zaba](https://translate.google.co.uk/?sl=pl&tl=en&text=Zaba&op=translate).

## Installation

You can install the development version of frogger from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("mrc-ide/frogger")
```

## Simulation model

The simulation model is implemented in a header-only C++ library located in [`inst/include/frogger.hpp`](inst/include/frogger.hpp). This location allows the C++ code to be imported in other R packages via specifying `LinkingTo: leapfrog` in the `DESCRIPTION` file.

The simulation model is callable in R via a wrapper function `run_model()` created with [Rcpp](https://www.rcpp.org).

You can control how the simulation model is run with the following arguments:

* `hiv_age_stratification` which must be "coarse" or "full". Coarse is run with 5-year age groups and full with single year ages.
* `run_child_model` which is `FALSE` by default. Set to `TRUE` to run the paediatric portion of the model.

## Example

The file `pjnz/bwa_aim-adult-art-no-special-elig_v6.13_2022-04-18.PJNZ` contains an example Spectrum file constructed
from default country data for Botswana with Spectrum (April 2022).

Prepare model inputs.

```{r example}
library(frogger)

pjnz <- system.file("pjnz/bwa_aim-adult-art-no-special-elig_v6.13_2022-04-18.PJNZ", 
                    package = "frogger", mustWork = TRUE)

demp <- prepare_leapfrog_demp(pjnz)
hivp <- prepare_leapfrog_projp(pjnz)
```

Simulate adult 'full' age group (single-year age) and 'coarse' age group (collapsed
age groups) models from 1970 to 2030 with 10 HIV time steps per year.

```{r simulate_leapfrog}
lsimF <- run_model(demp, hivp, 1970:2030, 10L, 
                   hiv_age_stratification = "full", run_child_model = FALSE)
lsimC <- run_model(demp, hivp, 1970:2030, 10L, 
                   hiv_age_stratification = "coarse", run_child_model = FALSE)
```

Compare the HIV prevalence age 15-49 years and AIDS deaths 50+ years. Deaths 50+ 
years are to show some noticeable divergence between the `"full"` and `"coarse"` 
age group simulations.

```{r sim_prev}
prevF <- colSums(lsimF$p_hiv_pop[16:50,,],,2) / colSums(lsimF$p_total_pop[16:50,,],,2)
prevC <- colSums(lsimC$p_hiv_pop[16:50,,],,2) / colSums(lsimC$p_total_pop[16:50,,],,2)

deathsF <- colSums(lsimF$p_hiv_deaths[51:81,,],,2)
deathsC <- colSums(lsimC$p_hiv_deaths[51:81,,],,2)

plot(1970:2030, prevF, type = "l", main = "Prevalence 15-49")
lines(1970:2030, prevC, col = 2)

plot(1970:2030, deathsF, type = "l", main = "AIDS Deaths 50+ years")
lines(1970:2030, deathsC, col = 2)
``` 

## Benchmarking

Install the package and then run the benchmarking script `./scripts/benchmark`

## lint

Lint R code with `lintr`

```R
lintr::lint_package()
```

Lint C++ code with [cpplint](https://github.com/cpplint/cpplint)

```console
cpplint inst/include/*
```

## Code design

### Simulation model

The simulation model is implemented as templated C++ code in `inst/include/frogger.hpp`.
This is so the simulation model may be developed as a standalone C++ library that can be called by other software without requiring R-specific code features.
The code uses header-only open source libraries to maximize portability.


### R functions

The file `src/frogger.cpp` contains R wrapper functions for the model simulation 
via [Rcpp](http://dirk.eddelbuettel.com/code/rcpp.html) and 
[RcppEigen](http://dirk.eddelbuettel.com/code/rcpp.eigen.html).


## Development notes

### Testing

There is some pre-prepared test data available to make tests run faster. This is generated and saved `./scripts/create_test_data.R`.

We also have some separate data written out in a generic format which can be read to test the model directly from C++. This is in `inst/standalone_model/data` in zipped files.

If this is your first time running you will need to unzip the standalone test data

```
./inst/standalone_model/extract_data
```

If you want to update the test data, it should be updated in the `./scripts/create_test_data.R` script so that we know how it was created and we can do it again fairly easily. Steps are
1. Update the script and generate the test data
1. Update the standalone data which is built from this `./scripts/update_standalone_data`. You might need to add a new mapping from R to serialized name if you are adding new input data
1. Unzip this for automated tests `./inst/standalone_model/extract_data`

### Simulation model

* The model was implemented using _Eigen::Tensor_ containers. These were preferred
  for several reasons:
   * Benchmarking found they were slighlty more efficient than _boost::multi_array_.
   * Column-major indexing in the same order as R
   * Other statistical packages (e.g. TMB, Stan) rely heavily on _Eigen_ so using _Eigen_ containers slims the dependencies.

### TODO

* Restructuring the model code to identify more common code
    * There are examples like general demographic projection and hiv population demographic projection which are running
      similar processes like ageing, non HIV mortality, migration. We should be able to write a function for e.g. ageing
      which we can run on each of our population matrices. Even for the HIV and ART stratified we can add overloaded
      function to work with higher dimension data
* Add a test that checks that no `double`s are used in `inst/include` dir. We should be using templated `real_type` for TMB
* Add a broad level overview of the algorithm - is there a diagram available?
* Convert string flags to the model to enums if we need to switch on them in several places. This should make it easier to reason about in C++ world and isolate the string checking to a single place
* Previously `hiv_negative_pop` was fixed size by having dimensions specified by template, how much does this speed up the code? Is there a better way to do this?
* Tidy up confusing looping see https://github.com/mrc-ide/frogger/pull/7#discussion_r1217847753
* Add R casting helpers which return better errors than Rcpp
  see https://github.com/mrc-ide/frogger/pull/7#discussion_r1217884684
* Add a helper to do 0 to base 1 conversion and check upper bounds
  see https://github.com/mrc-ide/frogger/pull/7#discussion_r1217888684
* Review what we pass as parameters - can some of these be computed in the struct ctor?
  e.g. https://github.com/mrc-ide/frogger/pull/7#discussion_r1217890466
* Refactor `OutputState` to take a struct of state-space dimensions instead of unpacking the subset of parameters we
  need. See https://github.com/mrc-ide/frogger/pull/12#discussion_r1245170775

## License

MIT © Imperial College of Science, Technology and Medicine

