---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  fig.cap = "",
  out.width = "100%"
)
```

```{r, include = FALSE}
## Build this README using `devtools::build_readme()`
```

# frogger R package

## Installation

Please install from our [r-universe](https://mrc-ide.r-universe.dev/builds):

```R
install.packages(
  "frogger",
  repos = c("https://mrc-ide.r-universe.dev", "https://cloud.r-project.org"))
```

You can install the development version of frogger from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("mrc-ide/frogger", subdir = "r-pacakge")
```

## Simulation model

The simulation model is implemented in a header-only C++ library located in [`r-package/inst/include/frogger.hpp`](r-package/inst/include/frogger.hpp). This location allows the C++ code to be imported in other R packages via specifying `LinkingTo: leapfrog` in the `DESCRIPTION` file.

```{=html}
> [!IMPORTANT]
> We use C++20 for this package. Please make sure you have a compiler that is compatible.
```

The simulation model is callable in R via a wrapper function `run_model()` created with [Rcpp](https://www.rcpp.org).

You can control how the simulation model is run by setting the `configuration` argument, see

```r
frogger::list_model_configurations()
```

for the list of available models. At the time of writing these are
- `DemographicProjection` - runs on the demographic projections
- `HivFullAgeStratification` - runs the demographic projection and HIV adult projection with single-year age groups
- `HivCoarseAgeStratification` - runs the demographic projection and HIV adult projection with 5-year age groups
- `ChildModel` - runs the demographic projection, single-year age stratified HIV adult projection and HIV child projection

## Example

The file `pjnz/bwa_aim-adult-art-no-special-elig_v6.13_2022-04-18.PJNZ` contains an example Spectrum file constructed
from default country data for Botswana with Spectrum (April 2022).

Prepare model inputs.

```{r example}
library(frogger)

pjnz <- system.file("pjnz/bwa_aim-adult-art-no-special-elig_v6.13_2022-04-18.PJNZ",
                    package = "frogger", mustWork = TRUE)

parameters <- prepare_leapfrog_parameters(pjnz)
```

Simulate adult 'full' age group (single-year age) model from 1970 to 2030 with 10 HIV time steps per year.

```{r simulate_leapfrog}
lsimF <- run_model(parameters, "HivFullAgeStratification", 1970:2030)
```

You can also simulate a model with 'coarse' age group (5-year age groups). You need to first prepare the coarse age group parameters

```{r simulate_coarse}
params_coarse <- prepare_leapfrog_parameters(pjnz, use_coarse_age_groups = TRUE)
lsimC <- run_model(params_coarse, "HivCoarseAgeStratification", 1970:2030)
```

Compare the HIV prevalence age 15-49 years and AIDS deaths 50+ years. Deaths 50+
years are to show some noticeable divergence between the `"full"` and `"coarse"`
age group simulations.

```{r sim_prev}
prevF <- colSums(lsimF$p_hivpop[16:50,,],,2) / colSums(lsimF$p_totpop[16:50,,],,2)
prevC <- colSums(lsimC$p_hivpop[16:50,,],,2) / colSums(lsimC$p_totpop[16:50,,],,2)

deathsF <- colSums(lsimF$p_hiv_deaths[51:81,,],,2)
deathsC <- colSums(lsimC$p_hiv_deaths[51:81,,],,2)

plot(1970:2030, prevF, type = "l", main = "Prevalence 15-49")
lines(1970:2030, prevC, col = 2)

plot(1970:2030, deathsF, type = "l", main = "AIDS Deaths 50+ years")
lines(1970:2030, deathsC, col = 2)
```

## Benchmarking

Install the package and then run the benchmarking script `./scripts/benchmark`

## lint

Lint R code with `lintr`

```R
lintr::lint_package()
```

Lint C++ code with [cpplint](https://github.com/cpplint/cpplint)

```console
cpplint r-package/inst/include/*
```

## Code design

### R functions

The file `src/frogger.cpp` contains R wrapper functions for the model simulation
via [Rcpp](http://dirk.eddelbuettel.com/code/rcpp.html) and
[RcppEigen](http://dirk.eddelbuettel.com/code/rcpp.eigen.html).

## Development notes

### Testing

There is some pre-prepared test data available to make tests run faster. This is generated and saved `./scripts/create_test_data.R`.

We also have some separate data written out in a generic format which can be read to test the model directly from C++. This is in `inst/standalone_model/data` in zipped files.

If this is your first time running you will need to unzip the standalone test data

```
./inst/standalone_model/extract_data
```

If you want to update the test data, it should be updated in the `./scripts/create_test_data.R` script so that we know how it was created and we can do it again fairly easily. Steps are
1. Update the script and generate the test data
1. Update the standalone data which is built from this `./scripts/update_standalone_data`. You might need to add a new mapping from R to serialized name if you are adding new input data
1. Unzip this for automated tests `./inst/standalone_model/extract_data`

## License

MIT © Imperial College of Science, Technology and Medicine

